Formalizing Stack Safety as a Security Property

### 2022-09-15 submission5/noexceptions.tex
For meeting

\vspace{\abovedisplayskip}
\begin{tabular}{| l | l | l |}
  \hline
  Set & Names & Purpose \\
  \hline
  CLR\_SV & \(rt_0 - rt_6\) & Caller-saved temps \\
  & \(ra_0 - ra_1\) & Caller-saved args / return vals \\
  & \(ra_2 - ra_7\) & Caller-saved args \\
  \hline
  CLE\_SV & \(rs_0 - rs_7\) & Callee-saved \\
  \hline
  PUBLIC & \(rra\) & Return Address \\
  & \(rsp\) & Stack Pointer \\
  & \(pc\) & Program Counter \\
  \hline
\end{tabular}
\vspace{\abovedisplayskip}

Collectively addresses and registers are {\em state elements} \(\component\)
in the set \(\COMPONENTS ::= \WORDS + \REGS\).
%
{\em Machine states} are drawn from a set \(\MACHS\) ranged over by \(\mach\)
that defines a mapping \(\mach[\component]\) from state elements to values.
The machine step relation
\(\mach \harpoonunder{\psi?} \mach' \in \MACHS \times (\psi+\bot) \times \MACHS\)
is left-total and may be labeled by a security operation, defined below.
While the tag-based enforcement mechanism may cause the machine to
failstop, we model this as the machine silently diverging by returning to the
same state. Since failstops are not distinguished from normal execution,
our properties will necessarily be {\it termination insensitive}.

For convenience, we will abbreviate multiple updates of a mapping
\(m[\component_0 \mapsto \word_0][\component_1 \mapsto \word_1]\dots\)
as \(m \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
are fomulae with free variables bound to sets in \(C\).

@@

Security operations are split into two categories: {\it privilege operations}
and {\it control operations}. Privilege operations are those that adjust the
privileges of future callees by allocating private or public data and by saving
catch points for possible exceptions. Allocations specify the offset from
the stack pointer and size of the object. Exceptions are tied to the address
of the buffer that will contain saved state.
%
\begin{align*}
  \phi \in \Phi ::= & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
  | & \mathbf{alloc\_pub} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
  | & \mathbf{setex} ~ \addr_{buf} \\
\end{align*}

Control operations are those that exchange
control between principals: calls (including tailcalls), returns, and thrown
exceptions. Calls and tailcalls identify the target of the call, and the argument registers.
[Still a todo: stack arguments.] Return needs no further information. Throwing
an exception specifies the address of the saved buffer.
%
\begin{align*}
  \chi \in X ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
  | & \mathbf{tailcall} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
  | & \mathbf{return} & \\
  | & \mathbf{throw} ~ \addr_{buf} \\
\end{align*}
%
A security operation may consist of any number of privilege operations and at most
one control operation.
%
\[\psi \in \Psi ::= \mathit{list ~ \Phi} | (X \times \mathit{list ~ \Phi})\]

Security principals are function activations, which we model as an {\em activation record}.
Such a record will track the status of the principal's data and the ways in which its callees
may return to it. To track the former, we define a set of {\it security classes}:
\[sc \in SEC ::= \sealed | \mathit{object} | \unsealed | \public\]

These indicate, from the perspective of a given principal, whether an element is
private to an inactive principal (\(\sealed\)),
in an allocated object that is accessible to the current principal (\(\mathit{object}\)),
available to be allocated (\(\unsealed\)), or
always accessible (\(\public\)).

A {\it security view} \(V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\)
assigns each element a security class.
The {\it initial view} \(V_0\) maps all stack locations and non-public registers
to \(\unsealed\) and all other locations and public registers to \(\public\).

An single principal might register multiple exception handlers over the
course of its execution, each of which will identify the value that the
program counter ought to have when an exception is thrown to it. We model
this as a partial map from addresses to addresses.
So an the activation record for the current activation is a pair of a view
and such a map:
\[(V,ex) \in \mathit{ACT} ::= \mathit{VIEW \times (\WORDS \rightharpoonup \WORDS)}\]

An inactive principal additional carries the address at which the program counter
should point in the event of a normal return.
\[(V, \addr_{ret}, ex) \in \mathit{INACT} ::= \mathit{VIEW \times \WORDS \times (\WORDS \rightharpoonup \WORDS)}\]

We represent a {\it call stack} as a pair of one active record and a list of inactive ones.
The initial call stack is \(\context_0 = ((V_0,[ ~ ]), [ ~ ])\).
\[(\alpha, \sigma) = \context \in \CONTEXTS ::= \mathit{ACT \times list ~ INACT}\]

@@

\judgmentthree[Control]
              {\(\mach \harpoonunder{\psi} \mach'\)}
              {\(\psi = (\chi, \overline{\phi})\)}
              {\(e = Ev ~ \chi ~ \mach\)}
              {\((\mach,\context) \stepstounder{e} (\mach',Op ~ \psi ~ \mach ~ \context)\)}

\judgment[NonControl]
         {\(\mach \harpoonunder{\psi} \mach'\)}
         {\((\mach,\context) \stepstounder{\tau} (\mach',Op ~ \psi ~ \mach ~ \context)\)}

\judgment[Default]
         {\(\mach \harpoonunder{\bot} \mach'\)}
         {\((\mach,\context) \stepstounder{\tau} (\mach',\context)\)}

\section{Context Updates With Calls, Returns, and Private Allocations}

In this section we will present the overlay steps for calls and returns.
They are given formally in Figure \ref{fig:callcontexts}.

In order to understand the motivation behind the call rule, though, we must first
describe how a function allocates objects. An allocation operation is parameterized
by an offset \(\mathit{off}\) and a size \(\mathit{sz}\). It will attempt to claim
a region of size \(\mathit{sz}\) starting at an offset of \(\mathit{off}\).
Claiming the region in this case means marking it as \(\mathit{object} ~ \mathbf{F}\),
where the boolean indicates whether the region is public.

\judgmenttwobrlong{\(b = \mach[sp] - \mathit{off}\)}
                  {\(\context = ((V, ex), \sigma)\)}
                  {\(V' = V \llbracket \addr \mapsto \mathit{object} ~ \mathbf{F} |
                    b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
                  {\(Pop ~ (\mathbf{alloc} ~ \mathit{off ~ sz}) ~ \context ~ \mach \triangleq
                    ((V',ex), \sigma)\)}

For calls, we define a function, {\it update view}, that defines the view
of a caller in terms of that of the callee. All argument registers are marked
\(\public\), along with all actual public registers. Callee-save registers,
including the stack pointer, are \(\sealed\); the callee is obligated to restore
them to their original value. The remaining caller-save registers are \(\unsealed\).
Then, all private objects are \(\sealed\), while the remainder of the stack retains
its old security context.

The formerly active principal's record is pushed onto the inactive list,
and replaced by a new record for the callee. Its return target is the intruction
immediately following that of the program counter before the call.
The callee's view is updated from the caller's with {\it update\_view}.

\judgmenttwo{\(\chi = \mathbf{call} ~ r_{target} ~ \overline{\reg_{args}}\)}
            {\(V' = \mathit{update\_view} ~ V ~ \overline{\reg_{args}}\)}
            {\(Cop ~ \chi ~ ((V,ex),\sigma) ~ \mach \triangleq
                ((V',[~]),(V, \mach[\PCname]+4, ex)::\sigma)\)}

@@

\judgmenttwo[Ret]
            {\(\chi = \mathbf{return}\)}
            {\(\sigma = (V,\addr_{ret},ex)::\sigma'\)}
            {\(Cop ~ \chi ~ (\_, \sigma) ~ \mach \triangleq
              ((V,ex), \sigma)\)}

\judgment[RetDef]
         {\(\chi = \mathbf{return}\)}
         {\(Cop ~ \chi ~ (\alpha, [~]) ~ \mach \triangleq
           (\alpha, [~])\)}

\begin{figure*}
  \[\mathit{reg\_view} ~ \overline{r_{args}} \triangleq \lambda \reg .
  \begin{cases}
    \public & \textnormal{if } \reg \in \overline{r_{args}} \textnormal{ or } \reg \in \mathit{PUBLIC} \\
    \unsealed & \textnormal{else if } \reg \in \mathit{CLR\_SV} \\
    \sealed & \textnormal{else} \\
  \end{cases}\]

  \[\mathit{update\_view} ~ V ~ \overline{r_{args}} \triangleq \lambda \component
  \begin{cases}
    \mathit{reg\_view} ~ \overline{r_{args}} ~ \component & \textnormal{if } \component \in \REGS \\
    \sealed & \textnormal{else if } V ~ \component = \mathit{object} \\

      V ~ \component & \textnormal{otherwise} \\
  \end{cases}\]

\caption{View updates for calls}
\label{fig:callcontexts}
\end{figure*}

@@

We define another relation that takes a trace until we have returned to a particular
activation, or past it (as might occur when we add exceptions.)
We write this \(d \downarrow (\mach,\context) \hookrightarrow \obsT\), where
\(d\) is the depth of the original call.

\judgment{\(|\sigma| \leq d\)}
         {\(d \downarrow (\mach,(\alpha,\sigma)) \hookrightarrow \tau\)}

\judgmenttwobrlong{\((\mach,(\alpha,\sigma)) \stepstounder{e} (\mach',\context')\)}
                  {\(|\sigma| > d\)}
                  {\(d \downarrow (\mach',\context') \hookrightarrow \obsT\)}
                  {\(d \downarrow (\mach,(\alpha,\sigma)) \hookrightarrow \notfinished{e}{\obsT}\)}

@@

We start by defining a second-order logical operator
\(d \uparrow P\), read ``\(P\) holds on return to depth \(d\),''
where \(P\) is a predicate on machine states. This is a coinductive relation
similar to ``weak until'' in temporal logic -- it holds if the program never
returns to depth \(d\).

\judgmenttwo[Returned]
            {\(|\sigma| \leq d }
            {\(P ~ (\mach,(\alpha,\sigma))\)}
            {\((d \uparrow P) ~ (\mach, (\alpha,\sigma))\)}

\judgmenttwobrlong[Step]
                  {\(|\sigma| > d\)}
                  {\((d \uparrow P) ~ (\mach', \context')\)}
                  {\((\mach, (\alpha,\sigma)) \stepstounder{e} (\mach', \context')\)}
                  {\((d \uparrow P) ~ (\mach, (\alpha,\sigma))\)}

Similarly, in {\it caller confidentiality}, we will want to compare future states,
so we give a binary equivalent, \(d \Uparrow R\), so that
\((\mach,\context) ~ (d \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
first states that return to depth \(d\) after \((\mach,\context)\) and \((\mach',\context')\).

\judgmenttwobrlong[Returned]
            {\(|\sigma_1| \leq d\)}
            {\(|\sigma_2| \leq d\)}
            {\((\mach_1,(\alpha_1,\sigma_1)) ~ R ~ (\mach_2,(\alpha_2,\sigma_2))\)}
            {\((\mach_1,(\alpha_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(\alpha_2,\sigma_2))\)}

\judgmenttwobrlong[Left]
              {\(|\sigma_1| > d\)}
              {\((\mach_1,(\alpha_1,\sigma_1)) \stepstounder{e} (\mach_1',\context_1')\)}
              {\((\mach_1',\context_1') ~ (d \Uparrow R) ~ (\mach_2,(\alpha_2,\sigma_2))\)}
              {\((\mach_1,(\alpha_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(\alpha_2,\sigma_2))\)}

\judgmenttwobrlong[Right]
              {\(|\sigma_2| > d\)}
              {\((\mach_2,(\alpha_2,\sigma_2)) \stepstounder{e} (\mach_2',\context_2')\)}
              {\((\mach_1,(\alpha_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,\context_2')\)}
              {\((\mach_1,(\alpha_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(\alpha_2,\sigma_2))\)}

@@

\definition Let \((\mach,(\alpha,T,V)::\context)\) be a state and
\(\components = \{\component | V ~ \component = \sealed\}\).
Let \(P\) be a predicate that holds on a state \((\mach',\context')\) if
\(\mach' ~ \parallel ~ \Delta(\mach,\mach') \cap \components\).
Then \(\mach\) enjoys {\it return-time integrity} if \(\context \downarrow P\) holds.

\definition A system enjoys {\it stack integrity} if, for all reachable states
\((\mach,\context)\) such that
\((\mach,\context) \stepstounder{\psi} (\mach',\context')\) with \(\psi = \mathbf{Call} ~ \_ ~ \_\),
\((\mach',\context')\) enjoys return-time integrity.

@@

\definition Let \((\mach,(\alpha,T,V)::\context)\) be a state and
\(\components = \{\component | V ~ \component \in \{\sealed, \unsealed\}\}\).

Then \((\mach,\context)\) enjoys {\it internal confidentiality} with respect to
some \(\context_{ret}\) if, for any \(\nach\)
that is a \(\components\)-variant of \(\mach\), we can take
\(\context_{ret} \uparrow (\mach,\context) \hookrightarrow \obsT\) and
\(\context_{ret} \uparrow (\nach,\context) \hookrightarrow \obsT'\) and have that
\(\obsT \simeq \obsT'\).

\definition Let \((\mach,(\alpha,T,V)::\context)\) be a state and
\(\components = \{\component | V ~ \component \in \{\sealed, \unsealed\}\}\).

Then \((\mach,\context)\) enjoys {\it return-time confidentiality} with respect to
some \(\context_{ret}\) if \((\mach,\context) ~ (\context_{ret} \Uparrow R) ~ (\nach,\context)\),
where \((\mach_1,\context_1) ~ R ~ (\nach_2,\context_2)\) if
\((\mach_1 \parallel \bar{\Diamond}(\mach,\mach_1,\nach,\nach_2)\).

\definition A system enjoys {\it caller confidentiality} if, for all reachable states
\((\mach,\context)\) such that \((\mach,\context) \stepstounder{\psi} (\mach',\context')\)
with \(\psi = \mathbf{Call} ~ \_ ~ \_\),
\((\mach',\context')\) enjoys both internal and return-time confidentiality
with respect to \(\context\).

@@

\definition Let \((\mach,(\alpha,T,V)::\context)\) be a state and
\(\components = \{\component | V ~ \component \in \{\sealed,\unsealed\}\}\).
Let \(P\) be a predicate that holds on a state \((\mach',\context')\) if
\(\mach' ~ \parallel ~ \Delta(\mach,\mach') \cap \components\).
Then \(\mach\) enjoys {\it post-return confidentiality} if \(\context \downarrow P\) holds.

\definition A system enjoys {\it callee confidentiality} if, for all reachable states
\((\mach,\context)\) such that \((\mach,\context) \stepstounder{\psi} (\mach',\context')\)
with \(\psi = \mathbf{Call} ~ \_ ~ \_\),
\((\mach',\context')\) enjoys post-return confidentiality
with respect to \(\context\).

@@

We can extend our basic machine to support public memory allocations that can
be accessed by anyone until they are deallocated on their owner's return.
We replace \(\mathbf{alloc} ~ \mathit{sizes}\) with \(\mathbf{alloc} ~ \mathit{objs}\),
where \(\mathit{objs}\) is a list of pairs of sizes and booleans, representing whether
the object in question should be public. The \(\mathit{update\_view}\)
function only seals private objects, and the remained maintain their \(\mathit{object}\)
class, making them exempt from both integrity and confidentiality properties.

\section{Exceptions}

We implement exceptions via another pair of macro-instructions:
\(\mathbf{setex} ~ a_{buf}\), akin to \(\mathit{setjmp}\), with the
jump buffer stored at \(a_{buf}\), and \(\mathbf{throwex} ~ a_{buf}\),
akin to \(\mathit{longjmp}\). The context updates for exceptions are given
in Figure \ref{fig:excontexts}. The update for \(\mathbf{setex}\) pushes
a \(\mathit{Catch}\) target onto the stack, identified by \(\addr_{buf}\),
with the same view as the current function. For \(\mathbf{throwex}\),
we pop the stack until we find a \(\mathit{Catch} ~ a_{buf} ~ \_ ~ \_\),
popping it as well. Rules apply in order.
[TODO: update these for the new \(\alpha\) stuff.]

\judgmenttwo[SetExC]
            {\(\psi = \mathbf{setex} ~ \addr_{buf}\)}
            {\(T' = \mathit{Catch} ~ \addr_{buf} ~ (\mach[pc] + 4) ~ \mach[sp]\)}
            {\(\mach, (T,V)::\context \constep{\psi} (T',V)::(T,V)::\context\)}

\judgment[ThrowExCCatch]
         {\(\psi = \mathbf{throwex} ~ \addr_{buf}\)}
         {\(\mach, (\mathit{Catch} \addr_{buf} ~ a_{pc} ~ a_{sp}, V)::\context \constep{\psi} \context\)}

\judgment[ThrowExCDone]
         {\(\psi = \mathbf{throwex} ~ \addr_{buf}\)}
         {\(m, [\mathit{Done},V] \constep{\psi} [\mathit{Done},V]\)}

\judgmenttwo[ThrowExCSkip]
         {\(\psi = \mathbf{throwex} ~ \addr_{buf}\)}
         {\(\mach, \context \constep{\psi} \context'\)}
         {\(\mach, (\_::\context) \constep{\psi} \context'\)}

\caption{Exception context updates}
\label{fig:excontexts}
\end{figure*}

@@

\judgmenttwo[TailCallC]
            {\(\psi = \mathbf{tailcall} ~ r_{target} ~ \overline{r_{args}}\)}
            {\(V' = \mathit{update\_view} ~ V ~ \overline{r_{args}}\)}
            {\(\mach, (\alpha,\mathit{Active},V)::\context \constep{\psi} (\alpha,\mathit{Active},V')::\context\)}

>>>>

\vspace{\abovedisplayskip}
\begin{tabular}{| l | l | l |}
  \hline
  Set & Names & Purpose \\
  \hline
  CLR\_SV & \(rt_0 - rt_6\) & Caller-saved temps \\
  & \(ra_0 - ra_1\) & Caller-saved args / return vals \\
  & \(ra_2 - ra_7\) & Caller-saved args \\
  \hline
  CLE\_SV & \(rs_0 - rs_7\) & Callee-saved \\
  & \(rra\) & Return Address \\
  & \(rsp\) & Stack Pointer \\  
  \hline
  PUBLIC & \(pc\) & Program Counter \\
  \hline
\end{tabular}
\vspace{\abovedisplayskip}

Collectively addresses and registers are {\em state elements} \(\component\)
in the set \(\COMPONENTS ::= \WORDS + \REGS\).
%
{\em Machine states} are drawn from a set \(\MACHS\) ranged over by \(\mach\)
that defines a mapping \(\mach[\component]\) from state elements to values.
The machine step function
\(\mach \xrightarrow{\psi} \mach' \in \MACHS \rightarrow \MACHS \times \Psi\)
may be labeled by a security operation, defined below.
While the tag-based enforcement mechanism may cause the machine to
failstop, we model this as the machine silently diverging by returning to the
same state. Since failstops are not distinguished from normal execution,
our properties will necessarily be {\it termination insensitive}.

For convenience, we will abbreviate multiple updates of a mapping
\(\mach[\component_0 \mapsto \word_0][\component_1 \mapsto \word_1]\dots\)
as \(\mach \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
are fomulae with free variables bound to sets in \(C\).

@@

Security operations are split into two categories: {\it privilege operations}
and {\it control operations}. Privilege operations are those that adjust the
privileges of future callees by allocating private or public data and by saving
catch points for possible exceptions. Allocations specify the offset from
the stack pointer and size of the object. Exceptions are tied to the address
of the buffer that will contain saved state.
%
\begin{align*}
  \phi \in \Phi ::= & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
  | & \mathbf{alloc\_pub} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
\end{align*}

Control operations are those that exchange
control between principals: calls (including tailcalls), returns, and thrown
exceptions. Calls and tailcalls identify the target of the call, and the argument registers.
[Still a todo: stack arguments.] Return needs no further information. Throwing
an exception specifies the address of the saved buffer.
%
\begin{align*}
  \chi \in \mathcal{X} ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
  | & \mathbf{tailcall} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
  | & \mathbf{return} & \\
\end{align*}
%
A security operation may consist of any number of privilege operations and at most
one control operation.
%
\[\psi \in \Psi ::= \mathit{list ~ \Phi} | (\mathcal{X} \times \mathit{list ~ \Phi})\]

Security principals are function activations: at any given time, the current
activation, and some number of pending ones. Each activation will have a {\it view}
of the system that maps each element to one of a set of {\it security classes}:
\[sc \in SEC ::= \sealed | \unsealed | \public\]
\[V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\]

These indicate, from the perspective of a given principal, whether an element is
private to an inactive principal (\(\sealed\)),
%in an allocated object that is accessible to the current principal (\(\mathit{object}\)),
available to be allocated (\(\unsealed\)), or
always accessible (\(\public\)).
The {\it initial view} \(V_0\) maps all stack locations and non-public registers
to \(\unsealed\) and all other locations and public registers to \(\public\).

A pending activation additional carries the addresses at which the program counter
and stack pointer should point in the event of a return.
We represent a {\it call stack} as a pair of the current activation's view
and a list of pending ones.
The initial call stack is \(\context_0 = (V_0, [ ~ ])\).
\[(V, \addr_{ret}, \addr_{sp}) \in \mathit{PEND} ::= \mathit{VIEW \times \WORDS \times \WORDS}\]
\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ PEND}\]

@@

\judgmentthree[Control]
              {\(\mach \xrightarrow{\psi} \mach'\)}
              {\(\psi = (\chi, \overline{\phi})\)}
              {\(e = Ev ~ \chi ~ \mach\)}
              {\((\mach,\context) \stepstounder{e} (\mach',Op ~ \psi ~ \context ~ \mach)\)}

\judgment[NonControl]
         {\(\mach \xrightarrow{\psi} \mach'\)}
         {\((\mach,\context) \stepstounder{\tau} (\mach',Op ~ \psi ~ \context ~ \mach)\)}

\section{Calls, Returns, and Private Allocations}

In this section we consider a simple setting with calls, returns, and private
allocations. Assume that \(\MACHS\) describe a RISC-V machine, where a transition
\(\mach \xrightarrow{\Psi?} \mach'\) is labeled as follows.
If \(\mach[\PCname]\) points to {\tt addi sp,sp,N}, \(\Psi? = \mathbf{alloc} ~ (0,N)\);
since we are not yet allocating public objects,
we don't need to distinguish objects from one another if they are allocated
simultaneously. Similarly, if \(\mach[\PCname]\) points to {\tt jal off ra}
and \(\mach'[\PCname] = a_f\) where \(a_f\) is the entry of a function
\(f\) that uses argument registers \(\overline{r}\), 
\(\Psi? = \mathbf{call} ~ f ~ \overline{r}\). And a step via the
instruction {\tt jalr ra} is labeled \(\mathbf{return}\).
This simplistic system serves to demonstrate our core properties.

Figure \ref{fig:example} gives a running example
of a function {\tt main} that takes a secret in its arguments and allocates
an array of potentially sensitive data. It then calls another function {\tt f},
and afterward may decide to {\tt publish} its secret based on the contents of
its sensitive data.

\begin{figure*}
  \begin{subfigure}[t]{.5\textwidth}
    {\tt
      void main(int argc, char*argv) \{

      ~ ~ int secret = argv[1];

      ~ ~ int sensitive[4] = \{0\};

      ~ ~ f();

      ~ ~ if (sensitive == 42) \{

      ~ ~ ~ ~ publish(secret);

      ~ ~ \}

      \}
    }
  \end{subfigure}
  \begin{subfigure}[t]{.5\textwidth}
    {\tt
      0: addi sp,sp,-20

      4: sw a1,16(sp)

      8: sd zero,8(sp)

      12: sd zero,0(sp)

      16: jal 84 ra

      20: lw a4,12(sp)

      24: li a5,42

      28: bne a4,a5,L

      32: ld a0,16(sp)

      36: jal 164 ra

      L(40): nop

      44: addi sp,sp,20

      48: return

    }
  \end{subfigure}

  \caption{Example: Main}
  \label{fig:example}
\end{figure*}

At the start of execution (PC 0) the initial call stack is \((V_0, [~])\).
Instruction 0 allocates 20 bytes starting at the stack pointer. This will have the
effect of marking those twenty bytes \(\sealed\), assuming they were previously
\(\unsealed\):

\judgmentbr{\(b = \mach[\SP] - \mathit{off}\)}
           {\(V' = V \llbracket \addr \mapsto \sealed |
             b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
           {\(Pop ~ (\mathbf{alloc} ~ \mathit{off, sz}) ~ (V,\sigma) ~ \mach \triangleq
             (V',\sigma)\)}

This memory being \(\sealed\) has no effect on the current activation --
the security guarantees apply at the start of the call. Our next security operation
occurs when we step from the call at instruction 16.

On a call, the formerly active principal's record is pushed onto the inactive list,
and replaced by a new record for the callee. Its return target is the instruction
immediately following that of the program counter before the call.
The callee's view is updated from the caller's such that the argument
registers are \(\public\) and any non-argument, caller-saved registers
are \(\unsealed\). All callee-save registers are always \(\sealed\), and
the program counter is always public, but this is already the case.

\[\begin{split}
\mathit{reg\_view} ~ V ~ \overline{r_{args}} \triangleq
V & \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR\_SV} \rrbracket \\
& \llbracket \reg \mapsto \public | \reg \in \overline{r_{args}} \rrbracket
\end{split}\]

\judgmenttwo{\(\chi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}}\)}
            {\(V' = \mathit{reg\_view} ~ V ~ \overline{\reg_{args}}\)}
            {\(Cop ~ \chi ~ (V,\sigma) ~ \mach \triangleq
              (V',(V, \mach[\PCname]+4, \mach[\SP])::\sigma)\)}

After the call step, we have a new activation, and {\tt main} is now pending.
{\tt main}'s pending record tracks its return address.
Everything that {\tt main} allocated, both {\tt secret} and {\tt sensitive},
{\tt f} views as \(\sealed\), along with its callee-saved registers.
Informally, the meaning of \(\sealed\) is: ``when control returns to
{\tt main}, this elements value should be the same as it is now.''
This is termed {\it return-time integrity}, and is defined below.
The remainder of the stack, and the caller-save registers, are \(\unsealed\).
The informal meaning of \(\unsealed\) is: ``{\tt f}'s behavior should not
depend on this element's value, but {\tt f} is free to overwrite it.''
This becomes relevant in out {\it confidentiality} properties.
Finally, \(\public\) elements are those that {\tt f} is allowed to
depend on, and may also overwrite.

@@

\judgmenttwo[Ret]
            {\(\chi = \mathbf{return}\)}
            {\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
            {\(Cop ~ \chi ~ (\_, \sigma) ~ \mach \triangleq
              (V, \sigma)\)}

\judgment[RetDef]
         {\(\chi = \mathbf{return}\)}
         {\(Cop ~ \chi ~ (V, [~]) ~ \mach \triangleq
           (V, [~])\)}

We are ready to define our first stack safety property for this system:
{\it well-bracketed control flow}.

\definition A system enjoys well-bracketed control flow if, for all states
\((\mach,(\_,(\_,\addr_{ret},\addr_{sp})) \stepstounder{\mathbf{return}} (\mach',\context)\),
\(\mach'[\PCname] = \addr_{ret}\) and \(\mach'[\SP] = \addr_{sp}\).

@@

We define another relation that takes a trace until we have returned from the
active principal.
We write this \(d \downarrow (\mach,\context) \hookrightarrow \obsT\), where
\(d\) is the depth of the current call.

\judgment{\(|\sigma| < d\)}
         {\(d \downarrow (\mach,(V,\sigma)) \hookrightarrow \tau\)}

\judgmenttwobrlong{\((\mach,(V,\sigma)) \stepstounder{e} (\mach',\context')\)}
                  {\(|\sigma| \geq d\)}
                  {\(d \downarrow (\mach',\context') \hookrightarrow \obsT\)}
                  {\(d \downarrow (\mach,(V,\sigma)) \hookrightarrow \notfinished{e}{\obsT}\)}

@@

We start by defining a second-order logical operator
\(d \uparrow P\), read ``\(P\) holds on return from depth \(d\),''
where \(P\) is a predicate on machine states. This is a coinductive relation
similar to ``weak until'' in temporal logic -- it holds if the program never
returns from depth \(d\).

\judgmenttwo[Returned]
            {\(|\sigma| < d\)}
            {\(P ~ (\mach,(V,\sigma))\)}
            {\((d \uparrow P) ~ (\mach, (V,\sigma))\)}

\judgmenttwobrlong[Step]
                  {\(|\sigma| \geq d\)}
                  {\((d \uparrow P) ~ (\mach', \context')\)}
                  {\((\mach, (V,\sigma)) \stepstounder{e} (\mach', \context')\)}
                  {\((d \uparrow P) ~ (\mach, (V,\sigma))\)}

Similarly, in {\it caller confidentiality}, we will want to compare future states,
so we give a binary equivalent, \(d \Uparrow R\), so that
\((\mach,\context) ~ (d \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
first states that return from depth \(d\) after \((\mach,\context)\) and \((\mach',\context')\).

\judgmenttwobrlong[Returned]
            {\(|\sigma_1| < d\)}
            {\(|\sigma_2| < d\)}
            {\((\mach_1,(V_1,\sigma_1)) ~ R ~ (\mach_2,(V_2,\sigma_2))\)}
            {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}

\judgmenttwobrlong[Left]
              {\(|\sigma_1| \geq d\)}
              {\((\mach_1,(V_1,\sigma_1)) \stepstounder{e} (\mach_1',\context_1')\)}
              {\((\mach_1',\context_1') ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}
              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}

\judgmenttwobrlong[Right]
              {\(|\sigma_2| \geq d\)}
              {\((\mach_2,(V_2,\sigma_2)) \stepstounder{e} (\mach_2',\context_2')\)}
              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,\context_2')\)}
              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}

@@

\definition Let \((\mach,\context)\) be a combined state with
\(\context = (V,\sigma)\), and
\(\components = \{\component | V ~ \component = \sealed\}\).
Let \(P\) be a predicate that holds on a state \((\mach',\context')\) if
\(\mach' ~ \parallel ~ \Delta(\mach,\mach') \cap \components\).
Then \(\mach\) enjoys {\it return-time integrity} if \(|\sigma| \downarrow P\) holds
on \((\mach,\context)\).

\definition A system enjoys {\it stack integrity} if, for all reachable states
\((\mach,\context)\) such that
\((\mach,\context) \stepstounder{e} (\mach',\context')\) with
\(e = \mathbf{callE} ~ \addr ~ \overline{\word}\),
\((\mach',\context')\) enjoys return-time integrity.

@@

\definition Let \((\mach,\context)\) be a state with \(\context = (V,\sigma)\), and
\(\components = \{\component | V ~ \component \in \{\sealed, \unsealed\}\}\).

Then \((\mach,\context)\) enjoys {\it internal confidentiality} with respect to
if, for any \(\nach\) that is a \(\components\)-variant of \(\mach\), we can take
\(|\sigma| \uparrow (\mach,\context) \hookrightarrow \obsT\) and
\(|\sigma| \uparrow (\nach,\context) \hookrightarrow \obsT'\) and have that
\(\obsT \simeq \obsT'\).

\definition Let \((\mach,\context)\) be a state with \(\context = (V,\sigma)\),
and \(\components = \{\component | V ~ \component \in \{\sealed, \unsealed\}\}\).
Then \((\mach,\context)\) enjoys {\it return-time confidentiality}
if, for any \(\nach\) that is a \(\components\)-variant of \(\mach\),
\((\mach,\context) ~ (|\sigma| \Uparrow R) ~ (\nach,\context)\),
where \((\mach_1,\context_1) ~ R ~ (\nach_2,\context_2)\) if
\((\mach_1,\context_1) \parallel \bar{\Diamond}(\mach,\mach_1,\nach,\nach_2)\).

\definition A system enjoys {\it caller confidentiality} if, for all reachable states
\((\mach,\context)\) such that \((\mach,\context) \stepstounder{e} (\mach',\context')\)
with \(e = \mathbf{callE} ~ \addr ~ \overline{\word}\),
\((\mach',\context')\) enjoys both internal and return-time confidentiality.

@@

\definition Let \((\mach,\context)\) be a state with \(\context = (V,\sigma)\), and
\(\components = \{\component | V ~ \component \in \{\sealed,\unsealed\}\}\).
Let \(P\) be a predicate that holds on a state \((\mach',\context')\) if
\(\mach' ~ \parallel ~ \Delta(\mach,\mach') \cap \components\).
Then \(\mach\) enjoys {\it post-return confidentiality} if \(|\sigma| \downarrow P\) holds.

\definition A system enjoys {\it callee confidentiality} if, for all reachable states
\((\mach,\context)\) such that \((\mach,\context) \stepstounder{e} (\mach',\context')\)
with \(\psi = \mathbf{callE} ~ \addr ~ \overline{\word}\),
\((\mach',\context')\) enjoys post-return confidentiality
with respect to \(\context\).

@@

We can extend our basic machine to support public memory allocations that can
be accessed by anyone until they are deallocated on their owner's return,
using the \(\mathbf{alloc\_pub}\) operation. This behaves identically to
\(\mathbf{alloc}\) except that it marks the locations \(\public\).
We should give an example of this, as it may also require us to use multiple
allocations simultaneously to interleave public and private allocations with a
single increase of the stack pointer.

@@

\judgmenttwo[TailCallC]
            {\(\chi = \mathbf{tailcall} ~ r_{target} ~ \overline{r_{args}}\)}
            {\(V' = \mathit{update\_view} ~ V ~ \overline{r_{args}}\)}
            {\(Cop ~ \chi ~ (V,\sigma) ~ (V',\sigma)\)}
